Field-informed Reinforcement Learning for Learning Large-Scale Collective Tasks

### 2023-05-18 paper.tex
abstract and intro ok

In this context, two main approaches have been proposed for creating distributed controllers out of macro-level task/goal descriptions: 
 \emph{manual design}, in which programmers build the controllers directly, and 
 \emph{automatic design}, which involves synthesizing programs using machine learning methods.
%
In this paper, we consider a new \emph{hybrid} approach called  \emph{Field-Informed reinforcement learning} (FIRL). We utilise manually designed \emph{computational fields} (globally distributed data structures) to manage the interaction between agents. Using \emph{Deep Q-learning} in combination with \emph{Graph Neural Networks} we enable the agents to learn the necessary local behaviour automatically to solve collective tasks. 
%
This allows us to create distributed controllers informed by a collective knowledge 
 that has been distilled during learning, but that uses only local information at runtime.
% 
We demonstrate the effectiveness of this new approach in simulated use cases 
 where tracking and covering tasks for swarm robotics are successfully solved. 
\end{abstract}

@@

A key question in designing swarm-like systems is how to create distributed controllers for the agents that enable them to perform complex collective tasks. 
 Two main approaches have been proposed for this problem: manual design and automatic design. 
%
In manual design, programmers build the controllers directly, 
 using domain knowledge and programming languages or frameworks that support distributed computation and communication like macro-programming approaches~\cite{DBLP:journals/corr/abs-2201-03473}. 
%
In automatic design, machine learning methods like \ac{MARL} and evolutionary algorithms are used to synthesize programs or policies for the agents from the high-level task or goal descriptions.
%
Both approaches have advantages and disadvantages. 
 Manual design allows programmers to specify desired properties for the system declaratively. However, manual design can be tedious and error-prone.
%
Automatic design can overcome these limitations by learning from data and experience. 
 However, automatic design can also suffer from several challenges, such as learning the right representation for agent communication.
%

In this paper, we propose a novel hybrid approach that combines manual and automatic design of distributed controllers.
This follows a novel trend in which high-level declarative programming languages are mixed with machine learning techniques in order to synthesise robust collective controllers~\cite{DBLP:conf/acsos/Aguzzi21,DBLP:conf/icdcs/AguzziCV22,DBLP:conf/coordination/AguzziCV22}.
Specifically, we propose a solution called \ac{FIRL} that utilises aggregate computing~\cite{Beal2015Computer} together with a \ac{GNN}~\cite{Zhou2020AIOpen} in combination with a reinforcement learning approach, namely \ac{DQN}~\cite{mnih2015human}. 
% First, aggregate computing is utilised to collect information from neighbours and create a local knowledge base for each agent. Second, we utilise a \ac{GNN} within the \ac{DQN}. 
 Here the \ac{GNN} is trained on field-information from aggregate computing and provides so-called node embeddings for each agent, serving as input for the \ac{DQN}. 
 The \ac{DQN} provides then the appropriate actions for the agent to achieve their tasks.
%We further combine this with reinforcement learning techniques to respond to changing conditions in the environment. 
%
The main contribution of \ac{FIRL} is that it allows us to create distributed controllers that are informed by a collective knowledge that has been distilled during training, but that use only local information when deployed. This way, \ac{FIRL} can achieve a balance between manual design and automatic design, combining the benefits of both approaches while mitigating their drawbacks.
%% Perhaps to much detail 
\ac{FIRL} can also be seen as a way of bridging the gap between symbolic and sub-symbolic AI methods, by integrating declarative programming with deep learning.

>>>>

In this context, two main approaches have been proposed for creating distributed controllers out of macro-level task/goal descriptions: 
 \emph{manual design}, in which programmers build the controllers directly, and 
 \emph{automatic design}, which involves synthesizing programs using machine learning methods.
%
In this paper, we consider a new \emph{hybrid} approach called  \emph{Field-Informed reinforcement learning} (FIRL). We utilise manually designed \emph{computational fields} (globally distributed data structures) to manage global agent coordination. Then, using \emph{Deep Q-learning} in combination with \emph{Graph Neural Networks} we enable the agents to learn the necessary local behaviour automatically to solve collective tasks, relying on those fields through local perception.
%
This allows us to create distributed controllers informed by a collective knowledge 
 that has been distilled during learning, but that uses only local information at runtime.
% 
We demonstrate the effectiveness of this new approach in simulated use cases 
 where tracking and covering tasks for swarm robotics are successfully solved. 
\end{abstract}

@@

A key question in designing swarm-like systems is how to create distributed controllers for the agents that enable them to perform complex collective tasks. 
 Two main approaches have been proposed for this problem: manual design and automatic design. 
%
In manual design, programmers build the controllers directly, 
 using domain knowledge and programming languages or frameworks that support distributed computation and communication like macro-programming approaches~\cite{DBLP:journals/corr/abs-2201-03473}. 
%
In automatic design, machine learning methods like \ac{MARL} and evolutionary algorithms are used to synthesize programs or policies for the agents from the high-level task or goal descriptions.
%
Both approaches have advantages and disadvantages.
 Manual design allows programmers to specify desired properties for the system declaratively, however, it can be tedious and error-prone.
%
Automatic design can overcome these limitations by learning from data and experience. 
 However, automatic design can also suffer from several challenges, such as learning the right representation for agent communication.
%

In this paper, we propose a novel hybrid approach that combines manual and automatic design of distributed controllers.
This follows a novel trend in which high-level declarative programming languages are mixed with machine learning techniques in order to synthesise robust collective controllers~\cite{DBLP:conf/acsos/Aguzzi21,DBLP:conf/icdcs/AguzziCV22,DBLP:conf/coordination/AguzziCV22}.
Specifically, we propose a solution called \ac{FIRL} that utilises aggregate computing~\cite{Beal2015Computer} together with a \ac{GNN}~\cite{Zhou2020AIOpen} in combination with a reinforcement learning approach, namely \ac{DQN}~\cite{mnih2015human}. 
% First, aggregate computing is utilised to collect information from neighbours and create a local knowledge base for each agent. Second, we utilise a \ac{GNN} within the \ac{DQN}. 
 Here the \ac{GNN} is trained on field-information from aggregate computing and provides so-called node embeddings for each agent, serving as input for the \ac{DQN}. 
 The \ac{DQN} provides then the appropriate actions for the agent to achieve their tasks.
%We further combine this with reinforcement learning techniques to respond to changing conditions in the environment. 
%
The main contribution of \ac{FIRL} is that it allows us to create distributed controllers that are informed by a collective knowledge that has been distilled during training, but that use only local information when deployed. This way, \ac{FIRL} can achieve a balance between manual design and automatic design, combining the benefits of both approaches while mitigating their drawbacks.
%% Perhaps to much detail 
\ac{FIRL} can also be seen as a way of bridging the gap between symbolic and sub-symbolic AI methods, by integrating declarative programming with deep learning.
%
Conceptually, \ac{FIRL} leverages the ideas of \emph{behaviour implicit communication} \cite{DBLP:journals/ijaci/CastelfranchiPT10} \cite{DBLP:conf/e4mas/TummoliniCRVO04}, whereby intelligent agents (here trained by DQN) achieve collective goals by learning how to use signs they left in the environment (here made of fields): much like ants collectively rely on pheromones they produce \cite{DBLP:journals/anor/Parunak97}.
